Stage-Splitting a Modal Language

### 2014-07-24 paper/semantics.tex
Fixed up more text to reflect new hatted vars.  Added pval fig.

The algorithm described above creates three different kinds of expressions which
cannot be evaluated further at a particular stage:
\begin{itemize}
\item 
Partial values ($\pvalsym$s) are stage \bbone\ terms that have been fully evaluated, 
but which may contain hatted variables. In the example above, 
\verb|(|$\hat {\tt y}$\verb|,7)|.
\item Residuals ($\ressym$es) are stage \bbtwo\ terms whose stage \bbone\
subexpressions have all been fully evaluated. In the example above,
\verb|(1+2)| and \verb|(let z = y*y*7 in z)|.
\item Values ($\valsym$s) are stage \bbtwo\ terms which are fully evaluated; these
are the results of a computation after both stages have been completed. In the
example above, the term evaluates to \verb|63|.
\end{itemize}
The $\redonesym$ judgment takes an open stage \bbone\ term to a {\em future environment} $\xi$ and a partial value $v$.
The future environment is a mapping from newly-created variables 
(which then may appear---hatted---in $v$)
to residuals.  It is the formal manifestation of the floating substitutions in our example.
For all of the normal features of \lang (\ref{fig:diaSemanticsCore}), first stage evaluation has the same behavior and effect on values as standard evaluation,
and the final future environment is gotten by merging the future environments of the subterms.

@@

At all of the normal features (\ref{fig:diaSemanticsSpec}), speculation does nothing but recursively speculate into every subexpression.
Once speculation finds a $\prev$ block, we resume stage \bbone\ evaluation of the contents, which produces a future context and (by canonical forms) a hatted variable ($\hat y$).
The context is then reified (using the $\reifysym$ judgment) into a series of let bindings, with $y$---stripped of its hat---at the bottom.

Because the inputs to evaluation and speculation are sometimes open on variables, we adorn our judgments with a context.
Since hatted variables can only be introduced to terms via substitution ({\em i.e.} they do not appear in the original program), 
this context is augmented only at stage \bbone\ substitution sites.
The second context ($\Gamma$) keeps track of stage \bbtwo\ variables, which are declared in original program.

>>>>

The algorithm described above creates three different kinds of expressions which
cannot be evaluated further at a particular stage:
\begin{itemize}
\item 
Partial values ($\pvalsym$s) are stage \bbone\ terms that have been fully evaluated, 
but which may contain hatted variables. In the example above, 
\verb|(|$\next \{\mathtt{\hat y}\}$\verb|,7)|.
\item Residuals ($\ressym$es) are stage \bbtwo\ terms whose stage \bbone\
subexpressions have all been fully evaluated. In the example above,
\verb|(1+2)| and \verb|(let z = y*y*7 in z)|.
\item Values ($\valsym$s) are stage \bbtwo\ terms which are fully evaluated; these
are the results of a computation after both stages have been completed. In the
example above, the term evaluates to \verb|63|.
\end{itemize}
The $\redonesym$ judgment takes an open stage \bbone\ term to a {\em future environment} $\xi$ and a partial value $v$.
The future environment is a mapping from newly-created variables 
(which then may appear---hatted---in $v$)
to residuals.  It is the formal manifestation of the floating substitutions in our example.
For all of the normal features of \lang (\ref{fig:diaSemanticsCore}), first stage evaluation has the same behavior and effect on values as standard evaluation,
and the final future environment is gotten by merging the future environments of the subterms.

@@

At all of the normal features (\ref{fig:diaSemanticsSpec}), speculation does nothing but recursively speculate into every subexpression.
Once speculation finds a $\prev$ block, we resume stage \bbone\ evaluation of the contents, which produces a future context and (by canonical forms) a $\next$-wrapped variable ($\next\{\hat y\}$).
The context is then reified (using the $\reifysym$ judgment) into a series of let bindings, with $\mathtt{\hat y}$---stripped of its $\next$---at the bottom.

The context ($\Gamma$) keeps track of stage \bbtwo\ variables in the input term. 
These both appear in the original program at stage \bbtwo\ and are inserted by the semantics.

As an optimization, we can include the special-case rule,
\begin{mathpar}
\inferdiaone [hat] {\red {\next~\hat y}{\cdot,\next~\hat y}}{\cdot}
\end{mathpar}
to avoid one-for-one variable bindings in the residual.
We used this implicitly in the example in the previous section.
